---
title: "Sperm size figures"
author: "jeremy"
date: "Wednesday, January 14, 2015"
output: pdf_document
---

Sperm size figures
========================================================

This is the output of an R markdown script - a fancy way of embedding r in a page so we can rerun/tune it as needed and then have the commands

I'm embedding, but hiding a bynch of stuff for readability - the actual code is in the .Rmd file.

```{r,echo=FALSE,hide=TRUE,message=FALSE}
library(phytools)
library(ape)
library(geiger)
library(dplyr)
library(ggplot2)
library(Hmisc)
library(picante)
library(adephylo)
library(caper)
library(ggtree)
library(psych)
```

Here's the tree from Kiontke et al 2011, figure 3
intial values I recieved had angaria and sp12 flipped - I've corrected it here
Here's the tree:

```{r,echo=FALSE}
tr<-read.tree(text="(C. sp. 1:0.356,(C. plicata:0.3,((C. guadeloupensis:0.223,((C. portoensis:0.158,C. virilis:0.201):0.005,((C. sp. 8:0.169,(C. angaria:0.031,C. castelli:0.037):0.131):0.073,(C. drosophilae:0.027,C. sp. 2:0.03):0.161):0.047):0.033):0.074,((C. kamaaina:0.14,(C. japonica:0.16,((C. imperialis:0.082,C. afra:0.182):0.009,((C. yunquensis:0.086,C. nouraguensis:0.085):0.038,C. macrosperma:0.159):0.057):0.012):0.059):0.041,(C. elegans:0.166,(((C. brenneri:0.129,C. doughertyi:0.111):0.042,(C. wallacei:0.083,C. tropicalis:0.07):0.033):0.013,(C. remanei:0.129,(C. sp. 5:0.127,(C. briggsae:0.047,C. nigoni:0.032):0.088):0.02):0.017):0.046):0.02):0.195):0.07):0.259);")

tr$tip.label<-c("C. sp. 1","C. plicata","C. guadeloupensis","C. portoensis","C. virilis","C. sp. 8","C. angaria","C. castelli","C. drosophilae","C. sp. 2","C. kamaaina","C. japonica","C. imperialis","C. afra","C. yunquensis","C. nouraguensis","C. macrosperma","C. elegans","C. brenneri","C. doughertyi","C. wallacei","C. tropicalis","C. remanei","C. sp. 5","C. briggsae","C. nigoni")

plot(tr)
```

Now read in the compiled data - we have oocytesize, spermsize, primaryspermatocytesize, and oocytearea. The species names in the files are inconsistent - they are fixed now. For now, the oocyte area is using the area formulation as made by the initial sperm size formula. All response variables are logged unless otherwise noted.

I have filtered out some of the data: Any oocytesize that is not stage -1, and any primaryspermatocyte data that is from hermaphrodites rather than males.

We now have clean data, clustered into species, strains or individuals. Variables ending in 1 are in species, 2 are strains and 3 are individuals. The non-clustered data is as is.


```{r,echo=FALSE}
setwd("C:/Users/jeremy/Desktop/geiger")
spermsizes<-read.table("compileddata.csv",header=T,sep=",")
oocytesize<-read.table("oocyteSizeCaeno.csv",header=T,sep=",")
primaryspermatocyte<-read.table("PrimarySpermatocyteSize.csv",header=T,sep=",")
bodysize<-read.csv("CaenoBodySize.csv")
eggsize<-read.csv("EmbryoSizeCaenoMarie_withSpp.csv")
#treat as ellipses
eggsize$area<-eggsize$Length.Mean*eggsize$Width.Mean*pi
#remove ones we don't have phylogenetic data for
eggsize<-eggsize[!eggsize$Species %in% c("latens","Ocheius tipulae","sinica"),]
#ugh, bodysize has a row with pasted in headers at line 1507 - removed it in my version of the file
#ugh, anagaria has "PS1010" and "PS1010 " replaced in my file
oocytesize<-oocytesize[oocytesize$oocyte.stage==-1,]
primaryspermatocyte<-primaryspermatocyte[primaryspermatocyte$Sex=="male",]
oocytesizearea<-read.csv("oocyteSizearea.csv")
oocytesizearea<-oocytesizearea[oocytesizearea$oocyte.stage==-1,]

weneednewnames <- list("brenneri"="C. brenneri",
                       "C. brenneri"="C. brenneri",
                       "C. angaria"="C. angaria",
                       "angaria"="C. angaria",
                       "angaria "="C. angaria",
                       "C. briggsae"="C. briggsae",
                       "briggsae"="C. briggsae",
                       "C. drosophilae"="C. drosophilae",
                       "drosophilae"="C. drosophilae",
                       "C. elegans"="C. elegans",
                       "elegans"="C. elegans",
                       "C. japonica"="C. japonica",
                       "japonica"="C. japonica",
                       "C. plicata"="C. plicata",
                       "plicata"="C. plicata",
                       "C. remanei"="C. remanei",
                       "remanei"="C. remanei",
                       "C. sp1"="C. sp. 1",
                       "C. sp 1"="C. sp. 1",
                       "sp1"="C. sp. 1",
                       "C. sp8"="C. sp. 8",
                       "C. sp 8"="C. sp. 8",
                       "sp8"="C. sp. 8",
                       "C. sp10"="C. doughertyi",
                       "C. sp. 10"="C. doughertyi",
                       "sp10"="C. doughertyi",
                       "doughertyi"="C. doughertyi",
                       "C. sp.11"="C. tropicalis",
                       "C. sp11"="C. tropicalis",
                       "C. sp. 11"="C. tropicalis",
                       "sp11"="C. tropicalis",
                       "tropicalis"="C. tropicalis",
                       "C. sp12"="C. castelli",
                       "C. sp. 12"="C. castelli",
                       "sp12"="C. castelli",
                       "castelli"="C. castelli",
                       "C. sp13"="C. virilis",
                       "C. sp. 13"="C. virilis",
                       "sp13"="C. virilis",
                       "virilis"="C. virilis",
                       "C. sp14"="C. imperialis",
                       "C. sp. 14"="C. imperialis",
                       "sp14"="C. imperialis",
                       "imperialis"="C. imperialis",
                       "C. sp15"="C. kamaaina",
                       "C. sp. 15"="C. kamaaina",
                       "sp15"="C. kamaaina",
                       "kamaaina"="C. kamaaina",
                       "C. sp16"="C. wallacei",
                       "C. sp. 16"="C. wallacei",
                       "sp16"="C. wallacei",
                       "wallacei"="C. wallacei",
                       "C. sp17"="C. nouraguensis",
                       "C. sp. 17"="C. nouraguensis",
                       "sp17"="C. nouraguensis",
                       "nouraguensis"="C. nouraguensis",
                       "C. sp 18"="C. macrosperma",
                       "C. sp18"="C. macrosperma",
                       "C. sp. 18"="C. macrosperma",
                       "sp18"="C. macrosperma",
                       "macrosperma"="C. macrosperma",
                       "C. sp19"="C. yunquensis",
                       "C. sp. 19"="C. yunquensis",
                       "sp19"="C. yunquensis",
                       "yunquensis"="C. yunquensis",
                       "C. sp2"="C. sp. 2",
                       "C. sp. 2"="C. sp. 2",
                       "sp2"="C. sp. 2",
                       "C. sp. 20"="C. guadeloupensis",
                       "sp20"="C. guadeloupensis",
                       "guadeloupensis"="C. guadeloupensis",
                       "C. sp5"="C. sp. 5",
                       "C. sp. 5"="C. sp. 5",
                       "sp5"="C. sp. 5",
                       "C. sp6"="C. portoensis",
                       "C. sp. 6"="C. portoensis",
                       "sp6"="C. portoensis",
                       "portoensis"="C. portoensis",
                       "C. sp7"="C. afra",
                       "C. sp. 7"="C. afra",
                       "sp7"="C. afra",
                       "afra"="C. afra",
                       "C. sp9"="C. nigoni",
                       "C. sp. 9"="C. nigoni",
                       "nigoni"="C. nigoni",
                       "sp9"="C. nigoni")

renameworms<-function(df){
  df$Species1<-as.character(df$Species)
  df$Species1<-weneednewnames[df$Species1]
  df$Species<-as.factor(unlist(df$Species1))
  df
}

bodysize<-renameworms(bodysize)
spermsizes<-renameworms(spermsizes)
oocytesize<-renameworms(oocytesize)
primaryspermatocyte<-renameworms(primaryspermatocyte)
oocytesizearea<-renameworms(oocytesizearea)
eggsize<-renameworms(eggsize)

clusterer<-function(df){
  x<-df %>% group_by(Species) %>%
    summarise(means=mean(area,na.rm=T),cv=sd(area,na.rm=T)/mean(area,na.rm=T))
  rownames(x) <- x$Species
  x
}

clusterer2<-function(df){
  x<-df %>% group_by(Species,Strain) %>%
    summarise(means=mean(area,na.rm=T),cv=sd(area,na.rm=T)/mean(area,na.rm=T))
  x
}

clusterer3<-function(df){
  x<-df %>% group_by(Species,Strain,individual) %>%
    summarise(means=mean(area,na.rm=T),cv=sd(area,na.rm=T)/mean(area,na.rm=T))
  x
}

spermsizes$area<-log(spermsizes$area)
spermsize1<-clusterer(spermsizes)
spermsize2<-clusterer2(spermsizes)
spermsize3<-clusterer3(spermsizes)

names(oocytesize)<-c("Species","Strain","individual","oocyte.stage","length.um","width.um","area")
oocytesize$area<-log(oocytesize$area)
oocytesize1<-clusterer(oocytesize)
oocytesize2<-clusterer2(oocytesize)
oocytesize3<-clusterer3(oocytesize)

names(primaryspermatocyte)<-c("Species","Strain","Sex","individual","area")
primaryspermatocyte$area<-log(primaryspermatocyte$area)
primaryspermatocyte1<-clusterer(primaryspermatocyte)
primaryspermatocyte2<-clusterer2(primaryspermatocyte)
primaryspermatocyte3<-clusterer3(primaryspermatocyte)

names(oocytesizearea)<-c("Species","Strain","individual","oocyte.stage","length.um","width.um","area")
oocytesizearea$area<-log(oocytesizearea$area)
oocytesizearea1<-clusterer(oocytesizearea)
oocytesizearea2<-clusterer2(oocytesizearea)
oocytesizearea3<-clusterer3(oocytesizearea)

bodysizelength<-bodysize[,1:4]
names(bodysizelength)<-c("Species","Strain","Sex","area")
bodysizelength$area<-log(bodysizelength$area)
bodysizelength1<-bodysizelength %>% group_by(Species,Strain,Sex) %>%
    summarise(means=mean(area,na.rm=T),cv=sd(area,na.rm=T)/mean(area,na.rm=T))


bodysizearea<-bodysize[,c(1:3,8)]
names(bodysizearea)<-c("Species","Strain","Sex","area")
bodysizearea$area<-log(bodysizearea$area)
bodysizearea1<-bodysizearea %>% group_by(Species,Strain,Sex) %>%
    summarise(means=mean(area,na.rm=T),cv=sd(area,na.rm=T)/mean(area,na.rm=T))


bodysizewidth<-bodysize[,c(1:3,7)]
names(bodysizewidth)<-c("Species","Strain","Sex","area")
bodysizewidth$area<-log(bodysizewidth$area)
bodysizewidth1<-bodysizewidth %>% group_by(Species,Strain,Sex) %>%
  summarise(means=mean(area,na.rm=T),cv=sd(area,na.rm=T)/mean(area,na.rm=T))


eggsize<-eggsize[,c(1,2,7)]
eggsize$area<-log(eggsize$area)
#we have only means - assume equal repeats
eggsize1<-eggsize %>% group_by(Species) %>%
    summarise(means=mean(area,na.rm=T),cv=sd(area,na.rm=T)/mean(area,na.rm=T))

plotpic<-function(df,tree,title,xlab,ylab){
  rownames(df)<-df$Species
  tree<-drop.tip(tree,tree$tip.label[!(tree$tip.label %in% df$Species)])
  df <- df[match(tree$tip.label,rownames(df)),]
  x1 <- pic(df$means, tree)
  y1 <- pic(df$cv, tree)
  plot(y1~x1,xlab=xlab,ylab=ylab,main=title,pch=19)
  fit1=lm(y1~x1)
  abline(fit1$coefficients,lwd=2,col=8)
  p=anova(fit1)$P[1]
  rp = vector('expression',1)
  rp[2]=substitute(expression(italic(p) == MYVALUE), list(MYVALUE = format(p,dig=3)))[2]
  legend("bottomright",bty="n",legend=rp)
}
```

###PIC plots - mean vs CV

Plotting the phylogenetic independent contrasts (PIC) of means of traits vs CV of traits. All means and CVs are grouped at strain level.

```{r,echo=FALSE}
plotpic(spermsize1,tr,"PIC of spermsize means vs CV","spermsize","CV")
plotpic(oocytesize1,tr,"PIC of oocytesize means vs CV","oocytesize","CV")
plotpic(primaryspermatocyte1,tr,"PIC of spermatocytesize means vs CV","spermatocytesize","CV")
plotpic(oocytesizearea1,tr,"PIC of oocytesizearea means vs CV","oocytesize","CV")
plotpic(bodysizearea1[bodysizearea1$Sex=="male",],tr,"PIC cv bodyvolume male","bodyvolume","CV")
plotpic(bodysizearea1[bodysizearea1$Sex=="female",],tr,"PIC cv bodyvolume female","bodyvolume","CV")
plotpic(bodysizelength1[bodysizearea1$Sex=="male",],tr,"PIC cv bodylength male","bodylength","CV")
plotpic(bodysizelength1[bodysizearea1$Sex=="female",],tr,"PIC cv bodylength female","bodylength","CV")
```

###table of all pairwise PIC correlations and p-values

tables with all pairwise PIC correlations. For the p-value table, those above diagonal are holm corrected

```{r, echo=FALSE}
tr1<-surface::nameNodes(tr)
totaldata<-as.data.frame(pic(spermsize1$means,tr1))
names(totaldata)<-c("spermsize")

totaldata$oocytesizearea<-pic(oocytesizearea1$means,tr1)
totaldata$oocytesize<-pic(oocytesize1$means,tr1)
totaldata$malebodyarea<-pic(bodysizearea1[bodysizearea1$Sex=="male",]$means,tr1)
totaldata$femalebodyarea<-pic(bodysizearea1[bodysizearea1$Sex=="female",]$means,tr1)
totaldata$malebodylength<-pic(bodysizelength1[bodysizelength1$Sex=="male",]$means,tr1)
totaldata$femalebodylength<-pic(bodysizelength1[bodysizelength1$Sex=="female",]$means,tr1)
totaldata$malebodywidth<-pic(bodysizewidth1[bodysizewidth1$Sex=="male",]$means,tr1)
totaldata$femalebodywidth<-pic(bodysizewidth1[bodysizewidth1$Sex=="female",]$means,tr1)

#both of the folliwng are very bad ideas. I am comparing the "same" nodes against each other
#but they are only joinging the same taxa, not the same depth or anything
#I need to think a little about what exactly I'm measuring
#and probably need to drop all taxa from both traits before a comparison
#so these cannot be done by corr.teste
data2<-merge(totaldata,pic(eggsize1$means,drop.tip(tr1,tr1$tip.label[!(tr1$tip.label %in% eggsize1$Species)])),by=0,all.x=T)
data3<-merge(totaldata,pic(primaryspermatocyte1$means,drop.tip(tr1,tr1$tip.label[!(tr1$tip.label %in% primaryspermatocyte1$Species)])),by=0,all.x=T)

totaldata$eggsize<-data2$y
totaldata$primaryspermatocyte<-data3$y
corr.test(totaldata)
```

###PIC plots - traits vs traits

```{r,echo=FALSE}
plotpic2<-function(df1,df2,tree,title,xlab,ylab){
  df<-merge(df1,df2,by="Species")
  tree<-drop.tip(tree,tree$tip.label[!(tree$tip.label %in% df$Species)])
  rownames(df)<-df$Species
  df <- df[match(tree$tip.label,rownames(df)),]
  x1 <- pic(df$means.x, tree)
  y1 <- pic(df$means.y, tree)
  plot(y1~x1,pch=19,main=title,xlab=xlab,ylab=ylab)
  fit1=lm(y1~x1)
  abline(fit1$coefficients,lwd=2,col=8)
  p=anova(fit1)$P[1]
  rp = vector('expression',1)
  rp[2]=substitute(expression(italic(p) == MYVALUE), list(MYVALUE = format(p,dig=3)))[2]
  legend("bottomright",bty="n",legend=rp)
}
plotpic2(spermsize1,oocytesize1,tr,"PIC of spermsize vs oocytesize","spermsize","oocytesize")
plotpic2(spermsize1,primaryspermatocyte1,tr,"PIC of spermsize vs spermatocytesize","spermsize","spermatocytesize")
plotpic2(primaryspermatocyte1,oocytesize1,tr,"PIC of spermatocytesize vs oocytesize","spermatocyte","oocytesize")
plotpic2(spermsize1,oocytesizearea1,tr,"PIC of spermsize vs oocytearea","spermsize","oocytearea")
plotpic2(primaryspermatocyte1,oocytesizearea1,tr,"PIC of spermatocytesize vs oocytearea","spermatocyte","oocytearea")
plotpic2(oocytesizearea1,oocytesize1,tr,"PIC of oocytesize vs oocytearea","oocytesize","oocytearea")
plotpic2(bodysizearea1[bodysizearea1$Sex=="male",],bodysizearea1[bodysizearea1$Sex!="male",],tr,"Body area male vs female","males","females")
plotpic2(bodysizearea1[bodysizearea1$Sex!="male",],oocytesize1,tr,"Body area female vs oocytesize","females","oocytes")
plotpic2(bodysizearea1[bodysizearea1$Sex=="male",],spermsize1,tr,"Body area male vs spermsize","males","sperms")
plotpic2(bodysizearea1[bodysizearea1$Sex=="male",],primaryspermatocyte1,tr,"Body area male vs spermatocytesize","males","spermatocytes")
plotpic2(bodysizelength1[bodysizelength1$Sex=="male",],bodysizelength1[bodysizelength1$Sex!="male",],tr,"Body length male vs female","males","females")
plotpic2(bodysizelength1[bodysizelength1$Sex!="male",],oocytesize1,tr,"Body length female vs oocytesize","females","oocytes")
plotpic2(bodysizelength1[bodysizelength1$Sex=="male",],spermsize1,tr,"Body length male vs spermsize","males","sperms")
plotpic2(bodysizelength1[bodysizelength1$Sex=="male",],primaryspermatocyte1,tr,"Body length male vs spermatocytesize","males","spermatocytes")
plotpic2(eggsize1,bodysizelength1[bodysizelength1$Sex!="male",],tr,"eggsize vs female length","egg size","length")
plotpic2(eggsize1,bodysizearea1[bodysizearea1$Sex!="male",],tr,"eggsize vs female volume","egg size","volume")
plotpic2(eggsize1,spermsize1,tr,"eggsize vs sperm size","egg size","sperm size")
plotpic2(eggsize1,oocytesize1,tr,"eggsize vs oocyte size","egg size","oocyte size")
```

###PGLS - testing for correlation with evolutionary models

Output are summary tables from fits. The OU model won't fit the spermatocyte fits due to overparameterisation.

```{r,echo=FALSE,warning=FALSE}
firsttwo<-merge(spermsize1,primaryspermatocyte1,by="Species",all=T)
secondtwo<-merge(oocytesizearea1,oocytesize1,by="Species",all=T)
fullpglsframe<-merge(firsttwo,secondtwo,by="Species",all=T)
names(fullpglsframe)<-c("Species","meansperm","cvsperm","meanprimsp","cvprimsp","meanooarea","cvooarea","meanoo","cvoo")
compsperm <- comparative.data(tr, fullpglsframe, Species, vcv=TRUE, vcv.dim=3)
mod1 <- pgls(meansperm ~ meanooarea, compsperm, lambda='ML')
mod2 <- pgls(meansperm ~ meanoo, compsperm, lambda='ML')
mod3 <- pgls(meansperm ~ meanprimsp, compsperm, lambda='ML')

summary(mod1)
summary(mod2)
summary(mod3)

rownames(fullpglsframe)<-fullpglsframe$Species
mod1ou <-gls(meansperm~meanooarea,correlation=corMartins(1,phy=tr),data=fullpglsframe)
mod2ou <-gls(meansperm~meanoo,correlation=corMartins(1,phy=tr),data=fullpglsframe)
#mod3ou <-gls(meansperm~meanprimsp,correlation=corMartins(1,phy=drop.tip(tr,tr$tip.label[!(tr$tip.label %in% primaryspermatocyte1$Species)])),data=fullpglsframe,na.action=na.omit)

summary(mod1ou)
summary(mod2ou)
```

Some phylogenetic signal methods - lambda

```{r,echo=FALSE,warning=FALSE}
fitlambda<-function(df,tr){
  holding<-df$mean
  names(holding)<-df$Species
  fitContinuous(tr,holding,model="lambda")
}

summary(fitlambda(spermsize1,tr))
summary(fitlambda(oocytesize1,tr))
summary(fitlambda(primaryspermatocyte1,tr))
summary(fitlambda(oocytesizearea1,tr))

```

Blombergs K
```{r,echo=FALSE,warning=FALSE}
fitblomberg<-function(df,tr){
  holding<-df$mean
  names(holding)<-df$Species
  phylosignal(holding,tr)
}

fitblomberg(spermsize1,tr)
fitblomberg(oocytesize1,tr)
fitblomberg(primaryspermatocyte1,tr)
fitblomberg(oocytesizearea1,tr)
```

Morans I
```{r,echo=FALSE,warning=FALSE}
fitmoran<-function(df,tr){
  tr<-drop.tip(tr,tr$tip.label[!(tr$tip.label %in% df$Species)])
  holding<-df$mean
  names(holding)<-df$Species
  abouheif.moran(holding,W=proxTips(tr))
}
fitmoran(spermsize1,tr)
fitmoran(oocytesize1,tr)
fitmoran(primaryspermatocyte1,tr)
fitmoran(oocytesizearea1,tr)
```

abouheifs cmean
```{r,echo=FALSE,warning=FALSE}
fitabouheif<-function(df,tr){
  tr<-drop.tip(tr,tr$tip.label[!(tr$tip.label %in% df$Species)])
  holding<-df$mean
  names(holding)<-df$Species
  abouheif.moran(holding,W=proxTips(tr),method="oriAbouheif")
}
fitabouheif(spermsize1,tr)
fitabouheif(oocytesize1,tr)
fitabouheif(primaryspermatocyte1,tr)
fitabouheif(oocytesizearea1,tr)
```

figure 3A with violin plots

```{r,echo=FALSE,warning=FALSE}
violinit<-function(f,specieslist){
  g=f[f$Species%in%specieslist,]
  p<-ggplot(g,aes(factor(Strain), area))
  p + facet_wrap(~Species,ncol = 3,scales = "free_x")+geom_violin(aes(fill = Species)) + geom_jitter(position =
    position_jitter(width = .2),alpha=0.5,)+theme(legend.position = "none") +
    ylab( expression(paste("area (", mu, m^{2},")")))+xlab("Strain")+
    theme(panel.margin = unit(0.05, "lines"))+theme(panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black")
  )
}

violinit(spermsizes,c("C. brenneri","C. remanei","C. sp. 8","C. macrosperma"))
violinit(oocytesize,c("C. brenneri","C. remanei","C. sp. 8","C. macrosperma"))
violinit(primaryspermatocyte,c("C. brenneri","C. remanei","C. sp. 8","C. macrosperma"))
violinit(oocytesizearea,c("C. brenneri","C. remanei","C. sp. 8","C. macrosperma"))
```

figure 3B

```{r,echo=FALSE,warning=FALSE}
clumper<-function(species,dfindv,dfstrain,dfspecies){
  holding=dfindv[dfindv$Species==species,]
  holding=rbind(holding,dfstrain[dfstrain$Species==species,])
  holding=rbind(holding,dfspecies[dfspecies$Species==species,])
  holding$plotter<-factor(holding$plotter,levels=c("intra individual","between individuals","between strains"))
  holding$plotcolour<-as.numeric(factor(holding$Strain))
  return(holding)
}

plotcvbysplits<-function(df,strainlist){
  dfstrain<-df%>% summarise(cv=sd(means)/mean(means),means=mean(means),plotter="between individuals")
  dfstrain$individual<-"dropped"
  dfspecies<-dfstrain %>% summarise(cv=sd(means)/mean(means),means=mean(means),plotter="between strains")
  dfspecies$individual="dropped"
  dfspecies$Strain="dropped"
  dfindv<-df
  dfindv$plotter<-"intra individual"
  x<-clumper(strainlist[1],dfindv,dfstrain,dfspecies)
  if(length(strainlist)>1){
    for(i in 2:length(strainlist)){
      x=rbind(x,clumper(strainlist[i],dfindv,dfstrain,dfspecies))
    }
  }
  p<-ggplot(x,aes(factor(plotter),cv))
  p+geom_jitter(aes(colour = factor(plotcolour),size = 5),position = position_jitter(width = .2))+facet_wrap(~Species,ncol = length(strainlist),scales = "free_x")+
    theme(panel.margin = unit(0.05, "lines"),legend.position = "none")+xlab("")+
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
          panel.background = element_blank(),axis.line = element_line(colour = "black"))
}


plotcvbysplits(primaryspermatocyte3,c("C. brenneri","C. remanei","C. sp. 8","C. macrosperma"))
plotcvbysplits(spermsize3,c("C. brenneri","C. remanei","C. sp. 8","C. macrosperma"))
plotcvbysplits(oocytesize3,c("C. brenneri","C. remanei","C. sp. 8","C. macrosperma"))
plotcvbysplits(oocytesizearea3,c("C. brenneri","C. remanei","C. sp. 8","C. macrosperma"))
```

figure mockup a
we are starting agin here, as I've modified some of the variables etc above
Easier to start with a clean slate
Clean and read in data:

```{r,echo=FALSE}
library(phytools)
library(ape)
library(geiger)
library(dplyr)
library(ggplot2)
setwd("C:/Users/jeremy/Desktop/geiger")
ff<-read.table("malevhermcomp.csv",header=T,sep=",")
ff<-renameworms(ff)
```

mean and sd

```{r}
groupedmeans=ff %>% group_by(Species,sex) %>% summarise(mean=mean(area,na.rm=T),n=n(),sd=sd(area,na.rm=T))
limits <- aes(ymax = mean + 1.96*sd/sqrt(n), ymin=mean - 1.96*sd/sqrt(n))
dodge <- position_dodge(width=0.9)
ggplot(data=groupedmeans, aes(x=Species, y=mean, fill=sex)) + geom_bar(stat="identity", position=position_dodge(), colour="black")+geom_errorbar(limits,position=dodge, width=0.25)+
	scale_fill_manual(values=c("gray","white"))+ylab( expression(paste("area (", mu, m^{2},")")))+xlab("Species")+coord_cartesian(ylim=c(0,25))+
	theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
     panel.background = element_blank(),axis.line = element_line(colour = "black"),legend.position = "none")
```

Figure mockup b
group to get the data I want
it's ugly, but it works

```{r}
groupedstrains=ff %>% group_by(Species,Strain) %>%
    do(data.frame(meanM=mean(.[which(.$sex=="M"),]$area,na.rm=T),meanH=mean(.[which(.$sex=="H"),]$area,na.rm=T),nM=sum(.$sex=="M"),nH=sum(.$sex=="H"),
                  sdM=sd(.[which(.$sex=="M"),]$area,na.rm=T),sdH=sd(.[which(.$sex=="H"),]$area,na.rm=T)))
groupedstrains$sem=1.96*groupedstrains$sdM/sqrt(groupedstrains$nM)
groupedstrains$seh=1.96*groupedstrains$sdH/sqrt(groupedstrains$nH)
p=ggplot(data=groupedstrains, aes(x=meanM, y=meanH, colour=Species, ymin = meanH - seh,ymax=meanH + seh,xmin = meanM - sem,xmax=meanM + sem))
p+geom_point(size=5)+coord_cartesian(ylim=c(0,15),xlim=c(0,33))+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
     panel.background = element_blank(),axis.line = element_line(colour = "black"),legend.position = "none")+ geom_errorbar(width=0.5)+geom_errorbarh(height=0.5)+
	 ylab(expression(paste("Hermaphrodite Sperm area (", mu, m^{2},")")))+xlab(expression(paste("Male Sperm area (", mu, m^{2},")")))+
	 geom_smooth(aes(group=Species), method="lm",size=1.5,colour="black")
```

figure1a - bars
need to rerun the above

```{r}
ff<-spermsizes<-read.table("C:/Users/jeremy/Desktop/geiger/compileddata.csv",header=T,sep=",")
ff<-renameworms(ff)
spermsize2<-ff %>% group_by(Species,Strain) %>% summarise(means=mean(area,na.rm=T),cv=sd(area,na.rm=T)/mean(area,na.rm=T),sem=mean(area,na.rm=T)/sqrt(n()))
countspec=function(dataframe){
	holding=c()
	for(i in 1:nrow(dataframe)){
		holding=c(holding,sum(dataframe$Species==dataframe[i,]$Species))
	}
	return(holding)
}
spermsize2$Species<-ordered(spermsize2$Species,levels=c(tr$tip.label))
spermsize2<-spermsize2[order(spermsize2$Species),]
spermsize2$Strain<-factor(spermsize2$Strain, levels = spermsize2$Strain)
spermsize2$specnum=countspec(spermsize2)
spermsize2$cols=cut(spermsize2$means,3)
findchange<-function(df){
    holding=c()
    for(i in 1:length(df$Species)){
        if(i==1){
            holding=c(holding,0)
        } else if(df$Species[i]==df$Species[i-1]){
            holding=c(holding,0)
        } else {
            holding=c(holding,1)
        }
    }
	return(holding)
}

spermsize2$change<-findchange(spermsize2)
w<-(1-(spermsize2$specnum-1)*0.05)/spermsize2$specnum
pos <- 0.5 * (cumsum(w) + cumsum(c(0, w[-length(w)])))

gapsizes<-function(df,pos){
	pos1<-pos
	for(i in 1:length(df$specnum)){
		if(df$specnum[i]==1){
			pos1[i:length(pos1)]<-pos1[i:length(pos1)]+0.1
		} else if(df$change[i]==1){
			pos1[i:length(pos1)]<-pos1[i:length(pos1)]+0.1
		} else {
			pos1[i:length(pos1)]<-pos1[i:length(pos1)]+0.05
		}
	}
	return(pos1)
}

pos<-gapsizes(spermsize2,pos)
ggplot(data=spermsize2,aes(x = pos, width = w, y = means, fill=cols,ymin=means-sem,ymax=means+sem)) + scale_fill_brewer()+coord_cartesian(ylim=c(0,280))+
    geom_bar(stat = "identity",colour="black") + scale_x_continuous(labels = spermsize2$Strain, breaks = pos)+geom_linerange()+
	theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
     panel.background = element_blank(),axis.line = element_line(colour = "black"),legend.position = "none")
```

large tree
first group all the data

```{r}
spermsize1<-ff %>% group_by(Species) %>% summarise(means=mean(area,na.rm=T))
spermsize2<-ff %>% group_by(Species,Strain) %>% summarise(means=mean(area,na.rm=T))
spermsize3<-ff %>% group_by(Species,Strain,individual) %>% summarise(means=mean(area,na.rm=T))
```

function to fix the format of spermsizes, and fit tree

```{r}
fitace <- function(spermsizes,tree){
  holding<-spermsizes$means
  names(holding)<-spermsizes$Species
  return(fastAnc(tree,holding,CI=T))
}
```

function to choose a subset - works on any level, just chooses one from each species

```{r}
subsetter<-function(spermsizes){
  x=lapply(split(spermsizes, spermsizes$Species),function(subdf) subdf[sample(1:nrow(subdf), 1),])
  do.call("rbind",x)
}
```

example usage fitace(subsetter(spermsizes2),tr)
now need to loop and keep the values

```{r}
looper<-function(spermsizes,reps,tree){
  outputdata=data.frame()
  outputCI=data.frame()
  for(i in 1:reps){
    outputdata=rbind(outputdata,fitace(subsetter(spermsizes),tree)$ace)
	outputCI=rbind(outputCI,fitace(subsetter(spermsizes),tree)$CI95)
  }
  names(outputdata)=seq(from=27,to=51)
  return(c(outputdata,outputCI))
}
```

ok we can call eg x=looper(spermsize3,10,tr)
then mean and sd, etc. etc.

```{r}
x<-fitace(spermsize1,tr)
AncSperm=x$ace
#or AncSperm=apply(x,2,mean) if it's bootstrapped
spermsize1$Species <- factor(spermsize1$Species, levels = c("C. sp. 1","C. plicata","C. guadeloupensis","C. portoensis","C. virilis","C. sp. 8","C. angaria","C. castelli","C. drosophilae","C. sp. 2","C. kamaaina","C. japonica","C. imperialis","C. afra","C. yunquensis","C. nouraguensis","C. macrosperma","C. elegans","C. brenneri","C. doughertyi","C. wallacei","C. tropicalis","C. remanei","C. sp. 5","C. briggsae","C. nigoni"))
spermsize=spermsize1[order(spermsize1$Species),]$means
names(spermsize)<-spermsize1[order(spermsize1$Species),]$Species
plotfancytree<-function(tree,nodesizes,terminalsizes){
  nodesizes<<-nodesizes
  terminalsizes<<-terminalsizes
  print(ggtree(tree)+geom_text(subset=.(isTip),aes(label=label),hjust=-0.19,fontface="italic")+
    geom_point(subset=.(!isTip),color="#1b9e77", size=sqrt(nodesizes/pi))+
    geom_point(subset=.(isTip),color="#7570b3", size=sqrt(terminalsizes/pi))+
    geom_text(subset=.(isTip),aes(label=round(terminalsizes,1)), hjust=1, vjust=-0.4, size=3)+
    geom_text(subset=.(!isTip),aes(label=round(nodesizes,1)), hjust=1.5, vjust=-0.4, size=3)+
    scale_x_continuous(expand = c(0.1, 0.1)))
  rm(list=c("terminalsizes","nodesizes"),pos=".GlobalEnv")
}
plotfancytree(tr,AncSperm,spermsize)
```


surface analysis
```{r}


library(surface)

tr<-read.tree(text="(C. sp. 1:0.356,(C. plicata:0.3,((C. guadeloupensis:0.223,((C. portoensis:0.158,C. virilis:0.201):0.005,((C. sp. 8:0.169,(C. angaria:0.031,C. castelli:0.037):0.131):0.073,(C. drosophilae:0.027,C. sp. 2:0.03):0.161):0.047):0.033):0.074,((C. kamaaina:0.14,(C. japonica:0.16,((C. imperialis:0.082,C. afra:0.182):0.009,((C. yunquensis:0.086,C. nouraguensis:0.085):0.038,C. macrosperma:0.159):0.057):0.012):0.059):0.041,(C. elegans:0.166,(((C. brenneri:0.129,C. doughertyi:0.111):0.042,(C. wallacei:0.083,C. tropicalis:0.07):0.033):0.013,(C. remanei:0.129,(C. sp. 5:0.127,(C. briggsae:0.047,C. nigoni:0.032):0.088):0.02):0.017):0.046):0.02):0.195):0.07):0.259);")

tr$tip.label<-c("C. sp. 1","C. plicata","C. guadeloupensis","C. portoensis","C. virilis","C. sp. 8","C. angaria","C. castelli","C. drosophilae","C. sp. 2","C. kamaaina","C. japonica","C. imperialis","C. afra","C. yunquensis","C. nouraguensis","C. macrosperma","C. elegans","C. brenneri","C. doughertyi","C. wallacei","C. tropicalis","C. remanei","C. sp. 5","C. briggsae","C. nigoni")
tr<-nameNodes(tr)
spermsizes<-read.table("compileddata.csv",header=T,sep=",")
spermsizes<-renameworms(spermsizes)
clusterer<-function(df){
  x<-df %>% group_by(Species) %>% 
    summarise(means=mean(area,na.rm=T),cv=sd(area,na.rm=T)/mean(area,na.rm=T))
  rownames(x) <- x$Species
  x
}
surfacesperm<-clusterer(spermsizes)

rownames(surfacesperm)<-surfacesperm$Species
surfacesperm <- surfacesperm[match(tr$tip.label,rownames(surfacesperm)),]
valuesin<-surfacesperm$means
names(valuesin)<-surfacesperm$Species
z<-runSurface(nameNodes(tr),as.data.frame(valuesin))

runonesim<-function(tr,fit,shifts){
  runSurface(tr,surfaceSimulate(tr, type="hansen-fit", hansenfit=fit, 
                  shifts=shifts, sample_optima=TRUE)$dat)
}


getvals<-function(tr,fit,shifts){
  sim<-runonesim(tr,fit,shifts)
  lengthofsim<-length(sim$bwd)
  sim$bwd[[lengthofsim]]$n_regimes
}

distributionofvals<-function(n,tr,fit,shifts){
  rdply(n,getvals(tr,fit,shifts),.progress = "text")
}

#increase/run this to bootstap - for now it takes ~30 secs per run
#distributionofvals(2,tr,z$bwd[[2]]$fit,z$bwd[[2]]$savedshifts)