---
title: "Sperm size figures"
author: "jeremy"
date: "Wednesday, January 14, 2015"
output: pdf_document
---

Sperm size figures
========================================================

This is the output of an R markdown script - a fancy way of embedding r in a page so we can rerun/tune it as needed and then have the commands

First we gotta load libraries:

```{r}
library(phytools)
library(ape)
library(geiger)
library(dplyr)
library(ggplot2)
library(Hmisc)
```

Here's the tree from Kiontke et al 2011, figure 3
intial values I recieved had angaria and sp12 flipped - I've corrected it here

```{r}
tr<-read.tree(text="(C. sp. 1:0.356,(C. plicata:0.3,((C. guadeloupensis:0.223,((C. portoensis:0.158,C. virilis:0.201):0.005,((C. sp. 8:0.169,(C. angaria:0.031,C. castelli:0.037):0.131):0.073,(C. drosophilae:0.027,C. sp. 2:0.03):0.161):0.047):0.033):0.074,((C. kamaaina:0.14,(C. japonica:0.16,((C. imperialis:0.082,C. afra:0.182):0.009,((C. yunquensis:0.086,C. nouraguensis:0.085):0.038,C. macrosperma:0.159):0.057):0.012):0.059):0.041,(C. elegans:0.166,(((C. brenneri:0.129,C. doughertyi:0.111):0.042,(C. wallacei:0.083,C. tropicalis:0.07):0.033):0.013,(C. remanei:0.129,(C. sp. 5:0.127,(C. briggsae:0.047,C. nigoni:0.032):0.088):0.02):0.017):0.046):0.02):0.195):0.07):0.259);")
```

The input read is a little ghetto - let's rename everthing now it's in

```{r}
tr$tip.label<-c("C. sp. 1","C. plicata","C. guadeloupensis","C. portoensis","C. virilis","C. sp. 8","C. angaria","C. castelli","C. drosophilae","C. sp. 2","C. kamaaina","C. japonica","C. imperialis","C. afra","C. yunquensis","C. nouraguensis","C. macrosperma","C. elegans","C. brenneri","C. doughertyi","C. wallacei","C. tropicalis","C. remanei","C. sp. 5","C. briggsae","C. nigoni")
```

And plot out the normal tree

```{r}
plot(tr)
```

Let's read in the compiled data, and apply a quick filter:
instructions say:"the most relevant measure would be the stage -1 and I would ignore the others."

```{r}
setwd("C:/Users/jeremy/Desktop/geiger")
spermsizes<-read.table("compileddata.csv",header=T,sep=",")
oocytesize<-read.table("oocyteSizeCaeno.csv",header=T,sep=",")
primaryspermatocyte<-read.table("PrimarySpermatocyteSize.csv",header=T,sep=",")
oocytesize<-oocytesize[oocytesize$oocyte.stage==-1,]
```

The given species names were a little off - let's fix them. The oocyte ones and spermatocyte ones are too, as well as inconsistent across the files. 
Let's fix it:

```{r}
weneednewnames <- list("brenneri"="C. brenneri",
                       "C. brenneri"="C. brenneri",
                       "C. angaria"="C. angaria",
                       "C. briggsae"="C. briggsae",
                       "C. drosophilae"="C. drosophilae",
                       "C. elegans"="C. elegans",
                       "C. japonica"="C. japonica",
                       "C. plicata"="C. plicata",
                       "C. remanei"="C. remanei",
                       "C. sp1"="C. sp. 1",
                       "C. sp 1"="C. sp. 1",
                       "C. sp8"="C. sp. 8",
                       "C. sp 8"="C. sp. 8",
                       "C. sp10"="C. doughertyi",
                       "C. sp. 10"="C. doughertyi",
                       "C. sp11"="C. tropicalis",
                       "C. sp. 11"="C. tropicalis",
                       "C. sp12"="C. castelli",
                       "C. sp. 12"="C. castelli",
                       "C. sp13"="C. virilis",
                       "C. sp. 13"="C. virilis",
                       "C. sp14"="C. imperialis",
                       "C. sp. 14"="C. imperialis",
                       "C. sp15"="C. kamaaina",
                       "C. sp. 15"="C. kamaaina",
                       "C. sp16"="C. wallacei",
                       "C. sp. 16"="C. wallacei",
                       "C. sp17"="C. nouraguensis",
                       "C. sp. 17"="C. nouraguensis",
                       "C. sp 18"="C. macrosperma",
                       "C. sp18"="C. macrosperma",
                       "C. sp. 18"="C. macrosperma",
                       "C. sp19"="C. yunquensis",
                       "C. sp. 19"="C. yunquensis",
                       "C. sp2"="C. sp. 2",
                       "C. sp. 2"="C. sp. 2",
                       "C. sp. 20"="C. guadeloupensis",
                       "C. sp5"="C. sp. 5",
                       "C. sp. 5"="C. sp. 5",
                       "C. sp6"="C. portoensis",
                       "C. sp. 6"="C. portoensis",
                       "C. sp7"="C. afra",
                       "C. sp. 7"="C. afra",
                       "C. sp9"="C. nigoni",
                       "C. sp. 9"="C. nigoni")

renameworms<-function(df){
  df$Species1<-as.character(df$Species)
  df$Species1<-weneednewnames[df$Species1]
  df$Species<-as.factor(unlist(df$Species1))
  df
}

spermsizes<-renameworms(spermsizes)
oocytesize<-renameworms(oocytesize)
primaryspermatocyte<-renameworms(primaryspermatocyte)
```

names should now be uniform.

Let's cluster stuff by mean and cv

```{r}
clusterer<-function(df){
  x<-df %>% group_by(Species) %>% 
    summarise(means=mean(area,na.rm=T),cv=sd(area,na.rm=T)/mean(area,na.rm=T))
  rownames(x) <- x$Species
  x
}

clusterer2<-function(df){
  x<-df %>% group_by(Species,Strain) %>% 
    summarise(means=mean(area,na.rm=T),cv=sd(area,na.rm=T)/mean(area,na.rm=T))
  x
}

clusterer3<-function(df){
  x<-df %>% group_by(Species,Strain,individual) %>% 
    summarise(means=mean(area,na.rm=T),cv=sd(area,na.rm=T)/mean(area,na.rm=T))
  x
}

spermsize1<-clusterer(spermsizes)
spermsize2<-clusterer2(spermsizes)
spermsize3<-clusterer3(spermsizes)

names(oocytesize)<-c("Species","Strain","individual","oocyte.stage","length.um","width.um","area")
oocytesize1<-clusterer(oocytesize)
oocytesize2<-clusterer2(oocytesize)
oocytesize3<-clusterer3(oocytesize)

names(primaryspermatocyte)<-c("Species","Strain","Sex","individual","area")
primaryspermatocyte1<-clusterer(primaryspermatocyte)
primaryspermatocyte2<-clusterer2(primaryspermatocyte)
primaryspermatocyte3<-clusterer3(primaryspermatocyte)

```

And now with PIC, figure 1D

```{r}
plotpic<-function(df,tree){
  tree<-drop.tip(tree,tree$tip.label[!(tree$tip.label %in% df$Species)])
  df <- df[match(tree$tip.label,rownames(df)),]
  x1 <- pic(df$means, tree)
  y1 <- pic(df$cv, tree)
  plot(y1~x1,xlab="PIC of Sperm Size",ylab="PIC of CV",pch=19)
  fit1=lm(y1~x1)
  abline(fit1$coefficients,lwd=2,col=8)
  p=anova(fit1)$P[1]
  rp = vector('expression',1)
  rp[2]=substitute(expression(italic(p) == MYVALUE), list(MYVALUE = format(p,dig=3)))[2]
  legend("bottomright",bty="n",legend=rp)
}

plotpic(spermsize1,tr)
plotpic(oocytesize1,tr)
plotpic(primaryspermatocyte1,tr)
```

Let's try some PIC of spermsize vs oocyte size!

```{r}
plotpic2<-function(df1,df2,tree){
  df<-merge(df1,df2,by="Species")
  tree<-drop.tip(tree,tree$tip.label[!(tree$tip.label %in% df$Species)])
  rownames(df)<-df$Species
  df <- df[match(tree$tip.label,rownames(df)),]
  x1 <- pic(df$means.x, tree)
  y1 <- pic(df$means.y, tree)
  plot(y1~x1,pch=19)
  fit1=lm(y1~x1)
  abline(fit1$coefficients,lwd=2,col=8)
  p=anova(fit1)$P[1]
  rp = vector('expression',1)
  rp[2]=substitute(expression(italic(p) == MYVALUE), list(MYVALUE = format(p,dig=3)))[2]
  legend("bottomright",bty="n",legend=rp)
}
plotpic2(spermsize1,oocytesize1,tr)
plotpic2(spermsize1,primaryspermatocyte1,tr)
plotpic2(primaryspermatocyte1,oocytesize1,tr)
```

Some phylogenetic signal methods - lambda

```{r}
fitlambda<-function(df,tr){
  holding<-df$mean
  names(holding)<-df$Species
  fitContinuous(tr,holding)
}

fitlambda(spermsize1,tr)
fitlambda(oocytesize1,tr)
fitlambda(primaryspermatocyte1,tr)

```

Blombergs K
```{r}
fitblomberg<-function(df,tr){
  holding<-df$mean
  names(holding)<-df$Species
  phylosignal(holding,tr)
}

fitblomberg(spermsize1,tr)
fitblomberg(oocytesize1,tr)
fitblomberg(primaryspermatocyte1,tr)
```



figure 3A with violin plots

```{r}
violinit<-function(f,specieslist){
  g=f[f$Species%in%specieslist,]
  p<-ggplot(g,aes(factor(Strain), area))
  p + facet_wrap(~Species,ncol = 3,scales = "free_x")+geom_violin(aes(fill = Species)) + geom_jitter(position =
    position_jitter(width = .2),alpha=0.5,)+theme(legend.position = "none") +
    ylab( expression(paste("area (", mu, m^{2},")")))+xlab("Strain")+
    theme(panel.margin = unit(0.05, "lines"))+theme(panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),panel.background = element_blank(), axis.line = element_line(colour = "black")
  )
}

violinit(spermsizes,c("C. brenneri","C. remanei","C. sp. 8","C. macrosperma"))
violinit(oocytesize,c("C. brenneri","C. remanei","C. sp. 8","C. macrosperma"))
violinit(primaryspermatocyte,c("C. brenneri","C. remanei","C. sp. 8","C. macrosperma"))
```

figure 3B

```{r}
spermsize2.1<-spermsize3 %>% summarise(cv=sd(means)/mean(means),means=mean(means),plotter="between individuals")
spermsize2.1$individual="dropped"
spermsize1.1<-spermsize2.1 %>% summarise(cv=sd(means)/mean(means),means=mean(means),plotter="between strains")
spermsize1.1$individual="dropped"
spermsize1.1$Strain="dropped"
spermsize3.1<-spermsize3
spermsize3.1$plotter<-"intra individual"
clumper<-function(species){
	holding=spermsize3.1[spermsize3.1$Species==species,]
	holding=rbind(holding,spermsize2.1[spermsize2.1$Species==species,])
	holding=rbind(holding,spermsize1.1[spermsize1.1$Species==species,])
	holding$plotter<-factor(holding$plotter,levels=c("intra individual","between individuals","between strains"))
  holding$plotcolour<-as.numeric(factor(holding$Strain))
	return(holding)
}

plotter<-function(strainlist){
		x<-clumper(strainlist[1])
		if(length(strainlist)>1){
			for(i in 2:length(strainlist)){
				x=rbind(x,clumper(strainlist[i]))
				}
			}
		p<-ggplot(x,aes(factor(plotter),cv))
		p+geom_jitter(aes(colour = factor(plotcolour),size = 5),position = position_jitter(width = .2))+facet_wrap(~Species,ncol = length(strainlist),scales = "free_x")+
    theme(panel.margin = unit(0.05, "lines"),legend.position = "none")+xlab("")+
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
       panel.background = element_blank(),axis.line = element_line(colour = "black"))
	}
plotter(c("C. brenneri","C. remanei","C. sp. 8","C. macrosperma"))
```

figure mockup a
we are starting agin here, as I've modified some of the variables etc above
Easier to start with a clean slate
Clean and read in data:

```{r}
library(phytools)
library(ape)
library(geiger)
library(dplyr)
library(ggplot2)
setwd("C:/Users/jeremy/Desktop/geiger")
ff<-read.table("malevhermcomp.csv",header=T,sep=",")
ff[ff$Species=="C. sp.11",]$Species="C. sp. 11"
```

mean and sd

```{r}
groupedmeans=ff %>% group_by(Species,sex) %>% summarise(mean=mean(area,na.rm=T),n=n(),sd=sd(area,na.rm=T))
limits <- aes(ymax = mean + 1.96*sd/sqrt(n), ymin=mean - 1.96*sd/sqrt(n))
dodge <- position_dodge(width=0.9)
ggplot(data=groupedmeans, aes(x=Species, y=mean, fill=sex)) + geom_bar(stat="identity", position=position_dodge(), colour="black")+geom_errorbar(limits,position=dodge, width=0.25)+
	scale_fill_manual(values=c("gray","white"))+ylab( expression(paste("area (", mu, m^{2},")")))+xlab("Species")+coord_cartesian(ylim=c(0,25))+
	theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
     panel.background = element_blank(),axis.line = element_line(colour = "black"),legend.position = "none")
```

Figure mockup b
group to get the data I want
it's ugly, but it works

```{r}
groupedstrains=ff %>% group_by(Species,Strain) %>%
    do(data.frame(meanM=mean(.[which(.$sex=="M"),]$area,na.rm=T),meanH=mean(.[which(.$sex=="H"),]$area,na.rm=T),nM=sum(.$sex=="M"),nH=sum(.$sex=="H"),
                  sdM=sd(.[which(.$sex=="M"),]$area,na.rm=T),sdH=sd(.[which(.$sex=="H"),]$area,na.rm=T)))
groupedstrains$sem=1.96*groupedstrains$sdM/sqrt(groupedstrains$nM)
groupedstrains$seh=1.96*groupedstrains$sdH/sqrt(groupedstrains$nH)
p=ggplot(data=groupedstrains, aes(x=meanM, y=meanH, colour=Species, ymin = meanH - seh,ymax=meanH + seh,xmin = meanM - sem,xmax=meanM + sem))
p+geom_point(size=5)+coord_cartesian(ylim=c(0,15),xlim=c(0,33))+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
     panel.background = element_blank(),axis.line = element_line(colour = "black"),legend.position = "none")+ geom_errorbar(width=0.5)+geom_errorbarh(height=0.5)+
	 ylab(expression(paste("Hermaphrodite Sperm area (", mu, m^{2},")")))+xlab(expression(paste("Male Sperm area (", mu, m^{2},")")))+
	 geom_smooth(aes(group=Species), method="lm",size=1.5,colour="black")
```

figure1a - bars
need to rerun the above

```{r}
spermsize2<-f %>% group_by(Species,Strain) %>% summarise(means=mean(area,na.rm=T),cv=sd(area,na.rm=T)/mean(area,na.rm=T),sem=mean(area,na.rm=T)/sqrt(n()))
countspec=function(dataframe){
	holding=c()
	for(i in 1:nrow(dataframe)){
		holding=c(holding,sum(dataframe$Species==dataframe[i,]$Species))
	}
	return(holding)
}
spermsize2$Species<-ordered(spermsize2$Species,levels=c(tr$tip.label))
spermsize2<-spermsize2[order(spermsize2$Species),]
spermsize2$Strain<-factor(spermsize2$Strain, levels = spermsize2$Strain)
spermsize2$specnum=countspec(spermsize2)
spermsize2$cols=cut(spermsize2$means,3)
findchange<-function(df){
    holding=c()
    for(i in 1:length(df$Species)){
        if(i==1){
            holding=c(holding,0)
        } else if(df$Species[i]==df$Species[i-1]){
            holding=c(holding,0)
        } else {
            holding=c(holding,1)
        }
    }
	return(holding)
}

spermsize2$change<-findchange(spermsize2)
w<-(1-(spermsize2$specnum-1)*0.05)/spermsize2$specnum
pos <- 0.5 * (cumsum(w) + cumsum(c(0, w[-length(w)])))

gapsizes<-function(df,pos){
	pos1<-pos
	for(i in 1:length(df$specnum)){
		if(df$specnum[i]==1){
			pos1[i:length(pos1)]<-pos1[i:length(pos1)]+0.1
		} else if(df$change[i]==1){
			pos1[i:length(pos1)]<-pos1[i:length(pos1)]+0.1
		} else {
			pos1[i:length(pos1)]<-pos1[i:length(pos1)]+0.05
		}
	}
	return(pos1)
}

pos<-gapsizes(spermsize2,pos)
ggplot(data=spermsize2,aes(x = pos, width = w, y = means, fill=cols,ymin=means-sem,ymax=means+sem)) + scale_fill_brewer()+coord_cartesian(ylim=c(0,280))+
    geom_bar(stat = "identity",colour="black") + scale_x_continuous(labels = spermsize2$Strain, breaks = pos)+geom_linerange()+
	theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
     panel.background = element_blank(),axis.line = element_line(colour = "black"),legend.position = "none")
```

large tree
first group all the data

```{r}
spermsize1<-f %>% group_by(Species) %>% summarise(means=mean(area,na.rm=T))
spermsize2<-f %>% group_by(Species,Strain) %>% summarise(means=mean(area,na.rm=T))
spermsize3<-f %>% group_by(Species,Strain,individual) %>% summarise(means=mean(area,na.rm=T))
```

function to fix the format of spermsizes, and fit tree

```{r}
fitace <- function(spermsizes,tree){
  holding<-spermsizes$means
  names(holding)<-spermsizes$Species
  return(fastAnc(tree,holding,CI=T))
}
```

function to choose a subset - works on any level, just chooses one from each species

```{r}
subsetter<-function(spermsizes){
  x=lapply(split(spermsizes, spermsizes$Species),function(subdf) subdf[sample(1:nrow(subdf), 1),])
  do.call("rbind",x)
}
```

example usage fitace(subsetter(spermsizes2),tr)
now need to loop and keep the values

```{r}
looper<-function(spermsizes,reps,tree){
  outputdata=data.frame()
  outputCI=data.frame()
  for(i in 1:reps){
    outputdata=rbind(outputdata,fitace(subsetter(spermsizes),tree)$ace)
	outputCI=rbind(outputCI,fitace(subsetter(spermsizes),tree)$CI95)
  }
  names(outputdata)=seq(from=27,to=51)
  return(c(outputdata,outputCI))
}
```

ok we can call eg x=looper(spermsize3,10,tr)
then mean and sd, etc. etc.

```{r}
x<-fitace(spermsize1,tr)
AncSperm=x$ace
#or AncSperm=apply(x,2,mean) if it's bootstrapped
spermsize1$Species <- factor(spermsize1$Species, levels = c("C. sp. 1","C. plicata","C. guadeloupensis","C. portoensis","C. virilis","C. sp. 8","C. angaria","C. castelli","C. drosophilae","C. sp. 2","C. kamaaina","C. japonica","C. imperialis","C. afra","C. yunquensis","C. nouraguensis","C. macrosperma","C. elegans","C. brenneri","C. doughertyi","C. wallacei","C. tropicalis","C. remanei","C. sp. 5","C. briggsae","C. nigoni"))
spermsize=spermsize1[order(spermsize1$Species),]$means
names(spermsize)<-spermsize1[order(spermsize1$Species),]$Species
plot(tr,label.offset=0,no.margin = TRUE,show.tip.label = F,x.lim=c(0,1),edge.width=1.5,edge.color=c(1))
#loads the graph into memory, then plots the anc. states on the nodes
#this is taken more or less from Nicos function
lastPP <- get("last_plot.phylo", envir = .PlotPhyloEnv)
  node <- (lastPP$Ntip + 1):length(lastPP$xx)
  XX <- lastPP$xx[node]
  YY <- lastPP$yy[node]
  for (i in 1:length(XX))
  {
    if (round(AncSperm[i], digits=1)==59.6)#this is a horrible hack to fix the labels if they are too big
    {
      points(XX[i],YY[i],col="red",lwd=round(AncSperm[i], digits=1)/max(round(spermsize,digit=1))*50,pch=19)
      text(XX[i]+0.02,YY[i],labels=round(AncSperm[i], digits=1),col="black",cex=1.2,adj=c(1,0))
    }
    else{
    points(XX[i],YY[i],col="red",lwd=round(AncSperm[i], digits=1)/max(round(spermsize,digit=1))*50,pch=19)
    text(XX[i],YY[i],labels=round(AncSperm[i], digits=1),col="black",cex=1.4,adj=c(1,0))
    }
  }
#now we label the tips
tip <- 1:lastPP$Ntip
XX <- lastPP$xx[tip]
YY <- lastPP$yy[tip]

points(XX,YY,col="blue",lwd=(round(spermsize,digit=1)/max(round(spermsize,digit=1))*50),pch=19)
text(XX+0.15,YY,labels=round(spermsize,digit=1),col="blue",cex=1.4)
text(XX+0.07,YY,labels=c("C. sp. 1","C. plicata","C. guadeloupensis","C. portoensis","C. virilis","C. sp. 8","C. angaria","C. castelli","C. drosophilae","C. sp. 2","C. kamaaina","C. japonica","C. imperialis","C. afra","C. yunquensis","C. nouraguensis","C. macrosperma", "C. elegans","C. brenneri","C. doughertyi","C. wallacei","C. tropicalis","C. remanei","C. sp. 5","C. briggsae","C. nigoni"),col="black",cex=1.2,font=3)
```
